Manuscript_Github_Pchabaudi_boilerplate

### 2022-08-04 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Analyse the stages comp within transmission heatmap

Which comparisons do I want to make?

```{r running_qusage}
ordered_stages <- ordering(unique(lifeCycle_labels),
         order_vector = times_stages_table$time)
#Remove sch.culture from the comparisons for now.
ordered_stages <- ordered_stages[!ordered_stages == 'sch.culture']
comparisons <- c(
  unlist(
    lapply(seq_along(ordered_stages),
           function(i){ 
             if(i < length(ordered_stages)){ 
               paste0(ordered_stages[i+1],'-',ordered_stages[i]) 
             } else {
               #When at the end, want it to cycle back round
               paste0(ordered_stages[1],'-',ordered_stages[i]) 
             } 
           }
    )
  )
)
```

@@

if(global_save_var){
  pdf(file = 'plots/stageTrans_logFoldheatmap.pdf',
      height = 8,
      width = 6)
  draw(ht)
  dev.off()
}

@@

>>>>

@@

if(global_save_var){
  pdf(file = 'plots/stageTrans_logFoldheatmap.pdf',
      height = 8,
      width = 6)
  draw(ht)
  dev.off()
}
```

Compare between stages within the same transmission method.

```{r qusage_transStage}
ordered_stages <- ordering(unique(lifeCycle_labels),
         order_vector = times_stages_table$time)
#Remove sch.culture from the comparisons for now.
ordered_stages <- ordered_stages[!ordered_stages == 'sch.culture']
comparisons_transStage.list <- lapply(c('rMT','SBP'), function(transmission){
  #Getting the comparisons across the time points for each transmission
  ordered_stages_trans <- str_subset(ordered_stages,
                               pattern = transmission)
  comparisons <- c(
    unlist(
      lapply(seq_along(ordered_stages_trans),
             function(i){ 
               if(i < length(ordered_stages_trans)){ 
                 paste0(ordered_stages_trans[i+1],'-',ordered_stages_trans[i]) 
               } else {
                 #When at the end, want it to cycle back round
                 paste0(ordered_stages_trans[1],'-',ordered_stages_trans[i]) 
               } 
             }
      )
    )
  )
})
names(comparisons_transStage.list) <- c('rMT', 'SBP')

# For the life cycle labels then extract the transmission + timepoint label
lifeCycle_labels <- str_extract(colnames(comp_stagesTrans.mat),
                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')

all_comparisons_transStage <- 
  lapply(c('rMT', 'SBP'), 
         function(transmission){
           #Extract the comparisons for this transmission method
           comparisons_transStage <- comparisons_transStage.list[[transmission]]
           #Subset the dataset based on the transmission and removing sch.cultures
           comp_transStage.mat <- chab_tpm_log2plus1.mat[
             ,str_detect(colnames(chab_tpm_log2plus1.mat), pattern = paste0(transmission,'_(?!sch.culture)'))
           ]
           #Get appropriate life cycle labels from this.
           lifeCycle_labels <- str_extract(colnames(comp_transStage.mat),
                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')
           results_transStage.list <- lapply(comparisons_transStage, 
                  function(comparison){
                    print(comparison)
                    list_comp <- lapply(
                      list(pir_geneset,
                           ShortLong_list,
                           subfam_list,
                           locus_list), 
                      function(gene_list){
                        print(deparse(substitute(gene_list)))
                        qs.results.subfam <- qusage(comp_transStage.mat,
                                                    lifeCycle_labels,
                                                    comparison,
                                                    gene_list,
                                                    n.points = 2^18)
                        list(results = qs.results.subfam, 
                             qstable = qsTable(qs.results.subfam) %>% 
                               mutate(contrast = qs.results.subfam$contrast))
                      })
                    names(list_comp) <- c('pir', 'shortlong', 'subfam', 'locus')
                    list_comp$combined <- bind_rows(lapply(list_comp, 
                                                           getElement, 
                                                           'qstable'))
                    return(list_comp)
                  })
           names(results_transStage.list) <- comparisons_transStage
           return(results_transStage.list)
         })
names(all_comparisons_transStage) <- c('rMT', 'SBP')

all_comparisons_transStage.df <- bind_rows(lapply(all_comparisons_transStage, 
                                           function(z) bind_rows(
                                             lapply(z, 
                                                    function(x) bind_rows(
                                                      lapply(x, 
                                                             function(y) y$qstable)))))) %>% 
  arrange(pathway.name)

@@

```{r heatmap_df_transStage}
transStages_ht.mat <- filter(all_comparisons_transStage.df, FDR < 0.05) %>% 
  mutate(pathway.name = factor(pathway.name, 
                               levels = ordering(pathway.name, 
                                                 order_vector = c('pirs', 
                                                                  'L', 
                                                                  'S', 
                                                                  'AAPL', 
                                                                  'ChAPL')))) %>% 
  arrange(pathway.name) %>% 
  melt(id.vars = c('contrast', 'pathway.name'),
       measure.vars  = c('log.fold.change')) %>% 
  dcast(contrast ~ pathway.name, 
        value.var = 'value') %>% 
  column_to_rownames('contrast') %>% 
  as.matrix

maxcol <- my_max(transStages_ht.mat)
#Multiply by minus one to use my_max to get the most negative value
mincol <- my_max(transStages_ht.mat*-1)*-1

#Transpose the matrix so that the stages (comparison) are on the x axis like other figures.
# Change the colnames of the transposed matrix so that they're just the times.
# Will split the heatmap by rMT and SBP when drawing the heatmap.
transStages_ht.mat.t <- t(transStages_ht.mat)
colnames(transStages_ht.mat.t) <- str_remove_all(colnames(transStages_ht.mat.t),
                                                 pattern = '(rMT|SBP)_')

ht <- Heatmap(transStages_ht.mat.t,
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        cluster_column_slices = FALSE,
        na_col = 'white',
        #Use factors to keep the ordering that we want.
        row_split = factor(str_extract(colnames(transStages_ht.mat),
                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'),
                              levels = unique(str_extract(colnames(transStages_ht.mat),
                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'))),
        row_order = colnames(transStages_ht.mat),
        row_gap = unit(.5,"cm"),
        row_names_gp = gpar(fontsize = 9),
        column_split = factor(str_extract(rownames(transStages_ht.mat),
                                   pattern = 'rMT|SBP'),
                              levels = unique(str_extract(rownames(transStages_ht.mat),
                                   pattern = 'rMT|SBP'))),
        column_gap = unit(.3, "cm"),
        border = TRUE,
        #Using this colour scheme because mincol is a lot less negative 
        # than maxcol is positive
        col = colorRamp2(c(mincol,
                           0,
                           maxcol/4,
                           maxcol/2,
                           maxcol),
                         c('blue', 'white', 'yellow', 'orange', 'red')),
        heatmap_legend_param = list(title = 'LogFoldChange'))
draw(ht)
```
